### 2023-12-31 tabsets.qmd
with50

---
title: "Tabsets"
format: 
  html:
    page-layout: article
#filters:
#  - webr
#format:
#  revealjs: 
#    smaller: true
#    toc-depth: 1
#    toc: true 
#    footer: <https://lotariohw26.github.io/ManifoldDestiny/>
#execute: 
#  cache: false
bibliography: man_bib.bib
---

```{r, echo=F, message=F, warning=F}
library(ManifoldDestiny)
library(ManifoldDestinyWASMP)
library(ManifoldDestinyWASMD)
library(RefManageR)
library(huxtable)
modlat <- eqpar$meql
md <- ManifoldDestinyWASMD::metad
source(paste0(rprojroot::find_rstudio_root_file(),'/script/simulations.R'))
source(paste0(rprojroot::find_rstudio_root_file(),'/script/applications.R'))
metad <- jsonlite::fromJSON(paste0(rprojroot::find_rstudio_root_file(),"/data-raw/metadata.json"))
mda <- jsonlite::fromJSON(paste0(rprojroot::find_rstudio_root_file(),"/data-raw/metadata.json"))
```

::: {.callout-caution collapse="true"}
## `r md$tabhead[1]`
```{r, echo=FALSE, results='asis',warning=F}
app_a_out <- app_n_out
app1_fil <- paste0(rprojroot::find_rstudio_root_file(), '/app_par.qmd')
text <- readLines(app1_fil)
cat(knitr::knit_child(text = paste(text[50:length(text)], collapse = "\n"), quiet = TRUE))
```
:::

::: {.callout-caution collapse="true"}
## `r md$tabhead[2]`
```{r, echo=FALSE, results='asis',warning=F}
app_a_out <- app_ex1_out 
app1_fil <- paste0(rprojroot::find_rstudio_root_file(), '/app_par.qmd')
text <- readLines(app1_fil)
cat(knitr::knit_child(text = paste(text[50:length(text)], collapse = "\n"), quiet = TRUE))
```
:::

::: {.callout-caution collapse="true"}
## `r md$tabhead[3]`
```{r, echo=FALSE, results='asis',warning=F}
app_a_out <- app_ex2_out 
app1_fil <- paste0(rprojroot::find_rstudio_root_file(), '/app_par.qmd')
text <- readLines(app1_fil)
cat(knitr::knit_child(text = paste(text[50:length(text)], collapse = "\n"), quiet = TRUE))
```
:::

::: {.callout-caution collapse="true"}
## `r md$tabhead[4]`
```{r, echo=FALSE, results='asis',warning=F}
app_a_out <- app_ex3_out 
app1_fil <- paste0(rprojroot::find_rstudio_root_file(), '/app_par.qmd')
text <- readLines(app1_fil)
cat(knitr::knit_child(text = paste(text[50:length(text)], collapse = "\n"), quiet = TRUE))
```
:::

::: {.callout-caution collapse="true"}
## Application 0 
```{r, echo=FALSE, results='asis',warning=F}
app_a_out <- app_0_out 
app1_fil <- paste0(rprojroot::find_rstudio_root_file(), '/app_par.qmd')
text <- readLines(app1_fil)
cat(knitr::knit_child(text = paste(text[50:length(text)], collapse = "\n"), quiet = TRUE))
```
:::

::: {.callout-caution collapse="true"}
## Application 1 
```{r, echo=FALSE, results='asis',warning=F}
app_a_out <- app_1_out 
app1_fil <- paste0(rprojroot::find_rstudio_root_file(), '/app_par.qmd')
text <- readLines(app1_fil)
cat(knitr::knit_child(text = paste(text[50:length(text)], collapse = "\n"), quiet = TRUE))
```
:::

::: {.callout-caution collapse="true"}
## Application 2 
```{r, echo=FALSE, results='asis',warning=F}
app_a_out <- app_2_out 
app1_fil <- paste0(rprojroot::find_rstudio_root_file(), '/app_par.qmd')
text <- readLines(app1_fil)
cat(knitr::knit_child(text = paste(text[50:length(text)], collapse = "\n"), quiet = TRUE))
```
:::
::: {.callout-caution collapse="true"}
## Application 3 
```{r, echo=FALSE, results='asis',warning=F}
app_a_out <- app_2_out 
app1_fil <- paste0(rprojroot::find_rstudio_root_file(), '/app_par.qmd')
text <- readLines(app1_fil)
cat(knitr::knit_child(text = paste(text[50:length(text)], collapse = "\n"), quiet = TRUE))
```
:::
::: {.callout-caution collapse="true"}
## Application 4 
```{r, echo=FALSE, results='asis',warning=F}
app_a_out <- app_2_out 
app1_fil <- paste0(rprojroot::find_rstudio_root_file(), '/app_par.qmd')
text <- readLines(app1_fil)
cat(knitr::knit_child(text = paste(text[50:length(text)], collapse = "\n"), quiet = TRUE))
```
:::
::: {.callout-caution collapse="true"}
## Application 5 
```{r, echo=FALSE, results='asis',warning=F}
app_a_out <- app_2_out 
app1_fil <- paste0(rprojroot::find_rstudio_root_file(), '/app_par.qmd')
text <- readLines(app1_fil)
cat(knitr::knit_child(text = paste(text[50:length(text)], collapse = "\n"), quiet = TRUE))
```
:::