### 2024-07-07 metaview.qmd
break

---
title: ""
output: html_document
---

```{r setup, include=FALSE}
library(yaml)
library(knitr)
## Load the YAML data from a file
yaml_data <- yaml::read_yaml("_variables.yml")
## Function to display each application's data
pretty_print_list <- function(x) {
  for (name in names(x)) {
    cat(paste0(name, ": "))
    value <- x[[name]]
    if (is.null(value)) {
      cat("NULL")
    } else if (is.atomic(value)) {
      cat(toString(value))
    } else if (is.list(value)) {
      cat("list()")
    }
    cat("\n")
  }
}

```

# App0

```{r, echo=F}
pretty_print_list(yaml_data$app0)
```