### 2024-07-08 paper.qmd
headlines

---
title: "{{< var titll >}}"
abstract: "Abstract goes here"
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# {{< var sec1.nmn >}}
## {{< var sec1.un1 >}}
## {{< var sec1.un2 >}}
## {{< var sec1.un3 >}}
## {{< var sec1.un4 >}}
## {{< var sec1.un5 >}}
## {{< var sec1.un6 >}}
## {{< var sec1.un7 >}}

# {{< var sec2.nmn >}}
## {{< var sec2.un1 >}}
## {{< var sec2.un2 >}}
## {{< var sec2.un3 >}}
## {{< var sec2.un4 >}}
## {{< var sec2.un4 >}}

# {{< var sec3.nmn >}}
## {{< var sec3.un1 >}}
## {{< var sec3.un2 >}}

# {{< var sec5.nmn >}}

# {{< var sec6.nmn >}}
## {{< var sec6.un1 >}}
## {{< var sec6.un2 >}}
## {{< var sec6.un3 >}}
## {{< var sec6.un4 >}}
## {{< var sec6.un4 >}}

# {{< var sec7.nmn >}}
## {{< var sec7.un1 >}}

# {{< var sec8.nmn >}}

```{r,echo=F}
knitr::knit_exit()
```

```{r, echo=FALSE, results='asis',warning=F}
#presm <- paste0(rprojroot::find_rstudio_root_file(), '/presentation.qmd')
#textm <- readLines(presm)
#cat(knitr::knit_child(text = paste(textm[25:length(textm)], collapse = "\n"), quiet = TRUE))
```