### 2024-07-09 app_par.qmd
tabsets

---
title: "Applicaton"
date: "`r Sys.Date()`"
#format: 
#  html:
#    page-layout: article
#format:
#  revealjs: 
#    smaller: true
#    toc-depth: 1
#    toc: true 
params:
  name: "0"
format: html
engine: knitr
filters:
  - webr
---

```{r, echo=F,message=FALSE,warning=FALSE}
ManifoldDestiny::wasmconload()
#library(ManifoldDestinyWASMP)
#library(ManifoldDestinyWASMD)
#library(ManifoldDestiny)
#library(dplyr)
#library(ggplot2)
#library(dplyr)
#library(combinat)
#library(htmltools)
#library(ggplot2)
#library(broom)
#library(combinat)
source(paste0(rprojroot::find_rstudio_root_file(),"/R/wasmconverting.R"))
source(paste0(rprojroot::find_rstudio_root_file(),"/R/wasmnonverting.R"))
mda <- app0[[2]]
app0_rp <- selreport(app0)
app_0_out <- seloutput(app0_rp)
mda$rcn
```

## Application: `r mda$nmn`
### Information
- Election system:
  - Type of race: `r mda$rcn` 
  - Candidates: `r mda$cnd` 
  - Modes of voting: `r mda$mod` 
- $H_0:$ Normal election 
- $H_1:$ Rigged election: 
  - Manifold object $`r mda$eq[1]`$
  - Solved for `r mda$va`

[Link to raw data and independent calculcations](`r mda$url`)

### Information

### {{< var tabh.h1 >}}

---

::: {.panel-tabset}

### {{< var tabs.h1 >}}

```{r, echo=F}
```
### {{< var tabs.h2 >}}
```{r, echo=F, out.width="60%"}
```
### {{< var tabs.h3 >}}
```{r, echo=F, out.width="60%"}
```
### {{< var tabs.h4 >}}
```{r, echo=F}
```
### {{< var tabs.h5 >}}
```{r, echo=F, out.width="60%"}
```
### {{< var tabs.h6 >}}
```{r, echo=F, out.width="60%"}
```
:::

---

```{r}
knitr::knit_exit()
```

---

::: {.panel-tabset}

### `r md$tabsets[[1]]`

```{r, echo=F}
kableExtra::kbl(app_a_out[['decs']]) %>% kableExtra::kable_paper(full_width = F) %>% kableExtra::kable_styling(font_size=8)
```
### `r md$tabsets[[2]]`
```{r, echo=F, out.width="60%"}
app_a_out[['corxy']]
```
### `r md$tabsets[[3]]`
```{r, echo=F, out.width="60%"}
app_a_out[['qunt']]
```
### `r md$tabsets[[4]]`
```{r, echo=F}
app_a_out[['ro3d']]
```
### `r md$tabsets[[5]]`
```{r, echo=F, out.width="60%"}
app_a_out[['regr']]
```
### `r md$tabsets[[6]]`
```{r, echo=F, out.width="60%"}
app_a_out[['resp']]
```
:::

---

### Ballot predictions based on limit information implied by the estimated Manifold object(s)

```{r, echo=F}
#app_a_out$cmp[,1] <- c("Mean total votes", "Number of precincts", "Zero prediction error", "In percentage", "Mean", "Standard Deviation", "Max ballot error", "1+/- prediction error", "2+/- prediction error", "2+/- prediction error", "Percentage within +/-3")
#kableExtra::kbl(app_a_out$cmp,row.names=F,col.names=c("Statistic","Value"),digits=2) %>% kableExtra::kable_paper(full_width = F) 
```