### 2024-07-09 tabsets.qmd

---
title: ""
---

```{r, echo=F, message=F, warning=F}
ManifoldDestiny::wasmconload()
md <- ManifoldDestiny::metd
source(paste0(rprojroot::find_rstudio_root_file(),'/script/R/simulations.R'))
source(paste0(rprojroot::find_rstudio_root_file(),'/script/R/applications.R'))
```

```{r, echo=F,message=FALSE,warning=FALSE}
ManifoldDestiny::wasmconload()
library(ManifoldDestiny)
source(paste0(rprojroot::find_rstudio_root_file(),"/R/wasmconverting.R"))
source(paste0(rprojroot::find_rstudio_root_file(),"/R/wasmnonverting.R"))
mda <- app0[[2]]
asrp <- selreport(app0)
asop <- seloutput(asrp)
```

::: {.callout-caution collapse="true"}
## {{< var tabh.h1 >}}
```{r, echo=FALSE, results='asis',warning=F}
#app_a_out <- app_0_out 
app1_fil <- paste0(rprojroot::find_rstudio_root_file(), '/app_par.qmd')
text <- readLines(app1_fil)
#cat(knitr::knit_child(text = paste(text[50:length(text)], collapse = "\n"), quiet = TRUE))
```
:::

::: {.callout-caution collapse="true"}
## {{< var tabh.h2 >}}
```{r, echo=FALSE, results='asis',warning=F}
#app_a_out <- app_1_out 
#app1_fil <- paste0(rprojroot::find_rstudio_root_file(), '/app_par.qmd')
#text <- readLines(app1_fil)
#cat(knitr::knit_child(text = paste(text[50:length(text)], collapse = "\n"), quiet = TRUE))
```
:::

::: {.callout-caution collapse="true"}
## {{< var tabh.h3 >}}
```{r, echo=FALSE, results='asis',warning=F}
#app_a_out <- app_2_out 
#app1_fil <- paste0(rprojroot::find_rstudio_root_file(), '/app_par.qmd')
#text <- readLines(app1_fil)
#cat(knitr::knit_child(text = paste(text[50:length(text)], collapse = "\n"), quiet = TRUE))
```
:::

::: {.callout-caution collapse="true"}
## {{< var sec6.un1 >}}
```{r, echo=FALSE, results='asis',warning=F}
#app_a_out <- app_2_out 
#app1_fil <- paste0(rprojroot::find_rstudio_root_file(), '/app_par.qmd')
#text <- readLines(app1_fil)
#cat(knitr::knit_child(text = paste(text[50:length(text)], collapse = "\n"), quiet = TRUE))
```
:::

::: {.callout-caution collapse="true"}
## {{< var sec6.un2 >}}
```{r, echo=FALSE, results='asis',warning=F}
#app_a_out <- app_2_out 
#app1_fil <- paste0(rprojroot::find_rstudio_root_file(), '/app_par.qmd')
#text <- readLines(app1_fil)
#cat(knitr::knit_child(text = paste(text[50:length(text)], collapse = "\n"), quiet = TRUE))
```
:::

::: {.callout-caution collapse="true"}
## {{< var sec6.un3 >}}
```{r, echo=FALSE, results='asis',warning=F}
#app_a_out <- app_2_out 
#app1_fil <- paste0(rprojroot::find_rstudio_root_file(), '/app_par.qmd')
#text <- readLines(app1_fil)
#cat(knitr::knit_child(text = paste(text[50:length(text)], collapse = "\n"), quiet = TRUE))
```
:::

::: {.callout-caution collapse="true"}
## {{< var sec6.un4 >}}
```{r, echo=FALSE, results='asis',warning=F}
#app_a_out <- app_2_out 
#app1_fil <- paste0(rprojroot::find_rstudio_root_file(), '/app_par.qmd')
#text <- readLines(app1_fil)
#cat(knitr::knit_child(text = paste(text[50:length(text)], collapse = "\n"), quiet = TRUE))
```
:::

```{r, echo=F}
knitr::knit_exit()
```