### 2024-07-09 abc.qmd
bbilio

---
title: ""
webr:
  show-startup-message: true      # Display status of webR initialization
  show-header-message: false      # Check to see if COOP&COEP headers are set for speed.
  packages: ['dplyr','data.table','readr','ggplotly']             # Pre-install dependencies
  cell-options:
    autorun: true
filters:
  - panelize
  - pyodide
  - webr
---

## Web-R

:::{.to-webr}
```{r}
1 + 1
```
:::

## Pyodide

:::{.to-pyodide}
```{python}
x = [1, 2]
print(x)
```
:::