### 2024-07-09 rigging.qmd
latest

---
title: ""
output: github_document
always_allow_html: true
execute: 
  cache: false
webr:
  show-startup-message: false      # Display status of webR initialization
  show-header-message: false      # Check to see if COOP&COEP headers are set for speed.
  packages: ['dplyr','data.table']             # Pre-install dependencies
filters:
  - panelize
  - webr
  - pyodide
---