### 2024-07-09 simulations.qmd

---
title: ""
output: github_document
always_allow_html: true
execute: 
  cache: false
webr:
  show-startup-message: false      # Display status of webR initialization
  show-header-message: false      # Check to see if COOP&COEP headers are set for speed.
  packages: ['dplyr','data.table','ggplot2','combinat','plotly','htmltools','broom'] # Pre-install dependencies
  cell-options:
    autorun: true
filters:
  - panelize
  - webr
  - pyodide
---

```{webr-r}
webr::install("ManifoldDestinyWASMP", repos = "https://lotariohw26.github.io/MD_WASMC")
webr::install("ManifoldDestinyWASMD", repos = "https://lotariohw26.github.io/MD_WASMC")
library(ManifoldDestinyWASMP)
library(ManifoldDestinyWASMD)
ls(package:ManifoldDestinyWASMP)
data(package='ManifoldDestinyWASMD')
library(ggplot2)
library(combinat)
library(plotly)
library(htmltools)
library(broom)
set.seed(1)
dfm <- (function(x){data.frame(P=seq(1,x),RV=as.integer(rnorm(x,100,30)))})(100)
probw <- c(m=0.51,s=0.10)
probva <- c(vdm=0.7,mdm=0.4,vds=0.10,mds=0.10)
probvb <- c(vdm=0.5,mdm=0.6,vds=0.10,mds=0.10)
ztech <- c(0,1)	
mdn <- list(frn=1,eq="alpha=k0+k1*x+k2*y")
app_bal <- list(df=ballcastsim(dfm,probw,probva,probvb,ztech),list(fr=1,eq="alpha=k0+k1*x+k2*y"))
app_bal[1]
```

```{webr-r}
```

**Last time updated: `r format(Sys.time(), "%Y-%m-%d %H:%M:%S")`**

```{r,echo=F}
knitr::knit_exit()
```