### 2024-07-09 index.qmd

---
title: "{{< var titll >}}"
output: github_document
execute: 
  cache: false
webr:
  show-startup-message: true      # Display status of webR initialization
  show-header-message: false      # Check to see if COOP&COEP headers are set for speed.
  packages: ['ggplot2','dplyr','data.table','readr','plotly','combinat','htmltools','broom','cowplot'] 
  cell-options:
    autorun: true
filters:
  - panelize
  - webr
---

```{r, echo=F}
# script/R/simulations.R
```

**Based upon the ideas developed by Edward K. Solomon**
![](pngs/md.png)

# Testing (latest from Ed's Youtube stream)

## Installling packages

```{webr-r}
#| echo: fenced
#| eval: true
webr::install("ManifoldDestinyWASMP", repos = "https://lotariohw26.github.io/MD_WASMC")
webr::install("ManifoldDestinyWASMD", repos = "https://lotariohw26.github.io/MD_WASMC")
library(ManifoldDestinyWASMP)
library(ManifoldDestinyWASMD)
ls(package:ManifoldDestinyWASMP)
data(package='ManifoldDestinyWASMD')
dfl <- ManifoldDestinyWASMD::yout
```

```{webr-r}
coi <- Countingprocess(dfl[[1]])
ges <- Estimation(coi$rdfc,2)
ges$regression("alpha=k0+k1*g+k2*h")
ges$diagnostics()
broom::tidy(ges$regsum[[1]])
broom::glance(ges$regsum[[1]])
```

[Nevada](https://docs.google.com/spreadsheets/d/1gCfmuK44sUeb0Hb0plAc8VllPSjtUiEHWlJjeHq3xXY/edit?gid=0#gid=0)

**Last time updated: `r format(Sys.time(), "%Y-%m-%d %H:%M:%S")`**

```{r, echo=F}
knitr::knit_exit()
```