### 2024-07-11 paper.qmd
abc

---
title: "{{< var titll >}}"
abstract: "Abstract goes here"
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

```{r}
library(ManifoldDestiny)
library(huxtable)
library(kableExtra)
source(paste0(rprojroot::find_rstudio_root_file(),'/script/R/simulations.R'))
```

# {{< var sec1.nmn >}}
## {{< var sec1.un1 >}}
## {{< var sec1.un2 >}}
## {{< var sec1.un3 >}}
## {{< var sec1.un4 >}}
## {{< var sec1.un5 >}}
## {{< var sec1.un6 >}}
## {{< var sec1.un7 >}}

# {{< var sec2.nmn >}}
## {{< var sec2.un1 >}}
## {{< var sec2.un2 >}}
## {{< var sec2.un3 >}}
## {{< var sec2.un4 >}}
## {{< var sec2.un5 >}}

# {{< var sec3.nmn >}}

## {{< var sec3.un1 >}}
```{r}
#| echo: false
#| output: false
apps <- basc
```
{{< include tab.qmd >}}

## {{< var sec3.un2 >}}
### Normal form
```{r}
#| echo: false
#| output: false
apps <- rigv[[1]]
```
{{< include tab.qmd >}}

### Hybrid
```{r}
#| echo: false
#| output: false
apps <- rigv[[2]]
```
{{< include tab.qmd >}}

### Opposition
```{r}
#| echo: false
#| output: false
apps <- rigv[[3]]
```
{{< include tab.qmd >}}

# {{< var sec5.nmn >}}

# {{< var sec6.nmn >}}

## {{< var sec6.un1 >}}
```{r}
#| echo: false
#| output: false
apps <- ManifoldDestiny::app1
```
{{< include tab.qmd >}}

## {{< var sec6.un2 >}}
```{r}
#| echo: false
#| output: false
apps <- ManifoldDestiny::app2
```
{{< include tab.qmd >}}

## {{< var sec6.un3 >}}
```{r}
#| echo: false
#| output: false
apps <- ManifoldDestiny::app2
```
{{< include tab.qmd >}}

## {{< var sec6.un4 >}}
```{r}
#| echo: false
#| output: false
apps <- ManifoldDestiny::app4
```
{{< include tab.qmd >}}

# {{< var sec7.nmn >}}
## {{< var sec7.un1 >}}

# {{< var sec8.nmn >}}

```{r,echo=F}
knitr::knit_exit()
```