### 2024-07-12 presentation.qmd
ac

---
title: "{{< var titll >}}"
subtitle: "Based upon the ideas developed by Edward K. Solomon"
format:
  revealjs: 
    smaller: true
    toc-depth: 1
    toc: true 
execute: 
  cache: true
footer: <https://lotariohw26.github.io/ManifoldDestiny/>
---

```{r, echo=F, message=F, warning=F, results='hide'}
ManifoldDestiny::wasmconload()
library(huxtable)
library(kableExtra)
library(ManifoldDestiny)
laws_df1 <- ManifoldDestiny::eqdef$meql
laws_df2 <- ManifoldDestiny::eqdef$meqs
source(paste0(rprojroot::find_rstudio_root_file(),'/script/R/stylized.R'))
#source(paste0(rprojroot::find_rstudio_root_file(),'/script/R/simulations.R'))
#source(paste0(rprojroot::find_rstudio_root_file(),'/script/R/applications.R'))
modlat <- eqpar$meql
```

# {{< var sec1.nmn >}}
Ed. Solomon

> "At the very core of this article lay the assumption of Causality, that the **Effect cannot precede the Cause**; likewise, the **Aggregate Percentage of a Candidate cannot precede the Election Day and the Mail-in Percentages of that candidate**.

## {{< var sec1.un1 >}}
- Two candidates: A, B
- County with three precincts: 1, 2, 3
- Two modes of voting
  - Number of ballots mode 1: S, T
  - Number of ballots mode 2: V, U

---

**Our two stylized numerical cases**

```{r, echo=F}
#| label: fig-1.1
#| fig-cap: "Case 1:.."
clr <- c("black","green","red","blue")
lego_hux_imp_i <- as_hux(towi[,1:3]) %>%
  set_text_color(5, 1:2, clr[1]) %>%
  set_background_color(1, 1:3, clr[2]) %>% 
  set_background_color(2, 1:3, clr[3]) %>%
  set_background_color(3, 1:3, clr[4]) %>%
  set_background_color(4, 2:3, clr[4]) %>%
  set_background_color(5, 3:3, clr[4]) %>%
  set_font_size(16)
lego_hux_imp_i
```

```{r, echo=F}
#| label: fig-1.2
#| fig-cap: "Case 2:..."
lego_hux_imp_ii <- as_hux(towii[,1:3]) %>%
  set_text_color(5, 1:2, clr[1]) %>%
  set_background_color(1, 1:3, clr[2]) %>% 
  set_background_color(2, 1:3, clr[3]) %>%
  set_background_color(3, 1:3, clr[4]) %>%
  set_background_color(4, 2:3, clr[4]) %>%
  set_background_color(5, 3:3, clr[4]) %>%
  set_font_size(16)
lego_hux_imp_ii
```

## {{< var sec1.un2 >}}
```{r, echo=F}
lego_hux_nor <- as_hux(matrix(c("S","V","T","U"), 2, 2)) %>%
  set_background_color(1, 1:2, "black") %>% 
  set_background_color(2, 1:2, "grey") %>% 
  set_text_color(1:2, 1:2, "red") %>% 
  set_all_borders(brdr(2, "solid", "black"))
lego_hux_nor %>% set_caption("")
```

Parameters:**$x$, $y$, $\zeta$, $\alpha$, $\lambda$**

\begin{equation*}
x=`r modlat$x_s[1]` \text{, }y=`r modlat$y_s[1]`
\end{equation*}
\begin{equation*}
\zeta=`r modlat$zeta_s[1]`
\end{equation*}
\begin{equation*}
\alpha=`r modlat$alpha_s[1]`=`r modlat$alpha_s[2]`
\end{equation*}
\begin{equation*}
\lambda=`r modlat$lamda_s[1]`=`r modlat$lamda_s[2]`
\end{equation*}

---

```{r, echo=F}
#| label: tbl-sfc1
#| tbl-cap: 'Election results applying standard form on case I'
#kableExtra::kbl(format(stfc1,digits=4)) %>% kableExtra::kable_classic_2(full_width = F) %>% kableExtra::row_spec(1, bold = F)
```

Numerical calculation of candidate one total percentage

- Precinct 1
  - $\alpha=\frac{`r cx[1]`+`r cy[1]`\cdot `r czeta[1]`}{1+`r czeta[1]`}=`r calpha[1]`$
- Precinct 2
  - $\alpha=\frac{`r cx[2]`+`r cy[2]`\cdot `r czeta[2]`}{1+`r czeta[2]`}=`r calpha[2]`$
- Precinct 3
  - $\alpha=\frac{`r cx[3]`+`r cy[3]`\cdot `r czeta[3]`}{1+`r czeta[3]`}=`r calpha[3]`$

---

```{r, echo=F}
#| label: tbl-sfc2
#| tbl-cap: 'Election results applying hybrid form on case II'
#kableExtra::kbl(format(stfc2,digits=4)) %>% kableExtra::kable_classic_2(full_width = F) %>% kableExtra::row_spec(1, bold = F)
```