### 2024-07-12 presentation.qmd
morepres

---
title: "{{< var titll >}}"
subtitle: "Based upon the ideas developed by Edward K. Solomon"
format:
  revealjs: 
    smaller: true
    toc-depth: 1
    toc: true 
execute: 
  cache: false
footer: <https://lotariohw26.github.io/ManifoldDestiny/>
---

```{r, echo=F, message=F, warning=F, results='hide'}
ManifoldDestiny::wasmconload()
library(huxtable)
library(kableExtra)
library(ManifoldDestiny)
laws_df1 <- ManifoldDestiny::eqdef$meql
laws_df2 <- ManifoldDestiny::eqdef$meqs
source(paste0(rprojroot::find_rstudio_root_file(),'/script/R/stylized.R'))
#source(paste0(rprojroot::find_rstudio_root_file(),'/script/R/simulations.R'))
#source(paste0(rprojroot::find_rstudio_root_file(),'/script/R/applications.R'))
modlat <- eqpar$meql
cx
```

# {{< var sec1.nmn >}}
Ed. Solomon

> "At the very core of this article lay the assumption of Causality, that the **Effect cannot precede the Cause**; likewise, the **Aggregate Percentage of a Candidate cannot precede the Election Day and the Mail-in Percentages of that candidate**.

## {{< var sec1.un1 >}}
- Two candidates: A, B
- County with three precincts: 1, 2, 3
- Two modes of voting
  - Number of ballots mode 1: S, T
  - Number of ballots mode 2: V, U

---

**Our two stylized numerical cases**

```{r, echo=F}
#| label: fig-1.1
#| fig-cap: "Case 1:.."
clr <- c("black","green","red","blue")
lego_hux_imp_i <- as_hux(towi[,1:3]) %>%
  set_text_color(5, 1:2, clr[1]) %>%
  set_background_color(1, 1:3, clr[2]) %>% 
  set_background_color(2, 1:3, clr[3]) %>%
  set_background_color(3, 1:3, clr[4]) %>%
  set_background_color(4, 2:3, clr[4]) %>%
  set_background_color(5, 3:3, clr[4]) %>%
  set_font_size(16)
lego_hux_imp_i
```

```{r, echo=F}
#| label: fig-1.2
#| fig-cap: "Case 2:..."
lego_hux_imp_ii <- as_hux(towii[,1:3]) %>%
  set_text_color(5, 1:2, clr[1]) %>%
  set_background_color(1, 1:3, clr[2]) %>% 
  set_background_color(2, 1:3, clr[3]) %>%
  set_background_color(3, 1:3, clr[4]) %>%
  set_background_color(4, 2:3, clr[4]) %>%
  set_background_color(5, 3:3, clr[4]) %>%
  set_font_size(16)
lego_hux_imp_ii
```

## {{< var sec1.un2 >}}
```{r, echo=F}
lego_hux_nor <- as_hux(matrix(c("S","V","T","U"), 2, 2)) %>%
  set_background_color(1, 1:2, "black") %>% 
  set_background_color(2, 1:2, "grey") %>% 
  set_text_color(1:2, 1:2, "red") %>% 
  set_all_borders(brdr(2, "solid", "black"))
lego_hux_nor %>% set_caption("")
```

Parameters:**$x$, $y$, $\zeta$, $\alpha$, $\lambda$**

\begin{equation*}
x=`r modlat$x_s[1]` \text{, }y=`r modlat$y_s[1]`
\end{equation*}
\begin{equation*}
\zeta=`r modlat$zeta_s[1]`
\end{equation*}
\begin{equation*}
\alpha=`r modlat$alpha_s[1]`=`r modlat$alpha_s[2]`
\end{equation*}
\begin{equation*}
\lambda=`r modlat$lamda_s[1]`=`r modlat$lamda_s[2]`
\end{equation*}

---

```{r, echo=F}
#| label: tbl-sfc1
#| tbl-cap: 'Election results applying standard form on case I'
kableExtra::kbl(format(stfc1,digits=4)) %>% kableExtra::kable_classic_2(full_width = F) %>% kableExtra::row_spec(1, bold = F)
```

Numerical calculation of candidate one total percentage

- Precinct 1
  - $\alpha=\frac{`r cx[1]`+`r cy[1]`\cdot `r czeta[1]`}{1+`r czeta[1]`}=`r calpha[1]`$
- Precinct 2
  - $\alpha=\frac{`r cx[2]`+`r cy[2]`\cdot `r czeta[2]`}{1+`r czeta[2]`}=`r calpha[2]`$
- Precinct 3
  - $\alpha=\frac{`r cx[3]`+`r cy[3]`\cdot `r czeta[3]`}{1+`r czeta[3]`}=`r calpha[3]`$

---

```{r, echo=F}
#| label: tbl-sfc2
#| tbl-cap: 'Election results applying hybrid form on case II'
kableExtra::kbl(format(stfc2,digits=4)) %>% kableExtra::kable_classic_2(full_width = F) %>% kableExtra::row_spec(1, bold = F)
```

## {{< var sec1.un3 >}}
```{r, echo=F}
lego_hux_imp <- as_hux(matrix(c("S","V","T","U"), 2, 2)) %>%
  set_background_color(1:2, 1, "black") %>% 
  set_background_color(1:2, 2, "grey") %>% 
  set_text_color(1:2, 1:2, "red") %>% 
  set_all_borders(brdr(2, "solid", "black"))
lego_hux_imp %>% set_caption("")
```

Parameters: $g$, $h$, $\Gamma$, $\alpha$, $\Omega$
\begin{equation*}
g=`r modlat$g_h[1]` \text{, } h=`r modlat$h_h[1]` 
\end{equation*}
\begin{equation*}
\Gamma=`r modlat$Gamma_h[1]`
\end{equation*}
\begin{equation*}
\alpha=`r modlat$alpha_h[1]`=`r modlat$alpha_h[2]`
\end{equation*}
\begin{equation*}
\Omega=`r modlat$Omega_h[1] `=`r modlat$Omega_h[2]`
\end{equation*}

---

```{r, echo=F}
#| label: tbl-hfc1
#| tbl-cap: 'Election results applying hybrid form case I'
kableExtra::kbl(format(hyfc1,digits=4)) %>% kableExtra::kable_classic_2(full_width = F) %>% kableExtra::row_spec(1, bold = F)
```

Numerical calculation of candidate one total percentage

- Precinct 1
  - $\alpha=\frac{`r cg[1]`+`r ch[1]`\cdot `r cGamma[1]`}{1+`r cGamma[1]`}=`r calpha[1]`$
- Precinct 2
  - $\alpha=\frac{`r cg[2]`+`r ch[2]`\cdot `r cGamma[2]`}{1+`r cGamma[2]`}=`r calpha[2]`$
- Precinct 3
  - $\alpha=\frac{`r cg[3]`+`r ch[3]`\cdot `r cGamma[3]`}{1+`r cGamma[3]`}=`r calpha[3]`$

---

```{r, echo=F}
#| label: tbl-hfc2
#| tbl-cap: 'Election results applying hybrid form case II'
kableExtra::kbl(format(hyfc2,digits=4)) %>% kableExtra::kable_classic_2(full_width = F) %>% kableExtra::row_spec(1, bold = F)
```

## {{< var sec1.un4 >}}
```{r, echo=F}
lego_hux_opo <- as_hux(matrix(c("S","T","U","V"), 2, 2)) %>%
  set_background_color(c(1),c(1), "black") %>% 
  set_background_color(c(2),c(2), "black") %>% 
  set_background_color(c(1),c(2), "grey") %>% 
  set_background_color(c(2),c(1), "grey") %>% 
  set_text_color(1:2, 1:2, "red") %>% 
  set_all_borders(brdr(2, "solid", "black"))
lego_hux_opo %>% set_caption("")
```

Parameters: $m$, $n$, $\xi$, $\lambda$, $\Omega$
\begin{equation*}
n=`r modlat$n_o[1]` \text{ , } m=`r modlat$m_o[1]`
\end{equation*}
\begin{equation*}
\xi=`r modlat$xi_o[1]`
\end{equation*}
\begin{equation*}
\Omega=`r modlat$Omega_o[1]`=`r modlat$Omega_o[2]`
\end{equation*}
\begin{equation*}
\lambda=`r modlat$lamda_o[1]`=`r modlat$lamda_o[2]`
\end{equation*}

---

```{r, echo=F}
#| label: tbl-ofc1
#| tbl-cap: 'Election results applying hybrid form case II'
kableExtra::kbl(format(opfc1,digits=4)) %>% kableExtra::kable_paper(full_width = F) %>% kableExtra::row_spec(1, bold = F)
```

---

```{r, echo=F}
#| label: tbl-ofc2
#| tbl-cap: 'Election results applying hybrid form case II'
kableExtra::kbl(format(opfc2,digits=4)) %>% kableExtra::kable_paper(full_width = F) %>% kableExtra::row_spec(1, bold = F)
```

## {{< var sec1.un5 >}}

- **Step 1:** Each registered voter assign themself (independent of voting technology) to either candidate A or candidate B
- **Step 2:** Voting technology realized
- **Step 3:** Registered voters determine whether to vote or not to vote 
- **Step 4:** If voting, do this by either mode 1 or by mode 2.

```{mermaid}
  flowchart LR
  A[Candidate A or B] --> R[A]
  R --> RV[Vote]
  RV --> RVE[Non-Mail]
  RV --> RVM[Mail]
  R --> RN[Not Vote]
  DV --> DVE[Non-Mail]
  DV --> DVM[Mail]
  A --> D[B]
  D --> DV[Vote]
  D --> Dn[Not Vote]
```

##

- Probability of assigning oneself to candidate A in precinct $i$
  - $p_{w,i} \sim N(\alpha_i,\sigma^2_{w,i})$
- Voting technology
  - $z_i \sim U(0,1)$
- Probability of not voting in precinct $i$
  - Candidate A, $p_{nz,a,i}=(1-z_i)p_{n,a,i}\sim N(\alpha_{a,n},\sigma^2_{a,n})$ <br>
  - Candidate B, $p_{nz,b,i}=(1-z_i)p_{n,b,i}\sim N(\alpha_{b,n},\sigma^2_{b,n})$
- Probability of voting by mode 2 in precinct $i$
  - Candidate A, $p_{m2,a,i}=(1-p_{nv,a,i})p_{a,mv}\sim N(\alpha_{a,mv},\sigma^2_{a,mv})$ 
  - Candidate B, $p_{m2,b,i}=(1-p_{nv,b,i})p_{b,mv}\sim N(\alpha_{a,mv},\sigma^2_{a,mv})$ 
- Probability of voting by mode 1 in precinct $i$
  - Candidate A, $p_{m1,a,i}=1-p_{m2,a,i}-p_{nz,a,i}$ 
  - Candidate B, $p_{m1,b,i}=1-p_{m2,b,i}-p_{nz,b,i}$

---

Neutral-voting assumption:

- Decission to vote or not to vote is independent of voting preferences:
  - $p_{nz,a,i}=p_{nz,a,i}$ 
- If voting, decission to vote by mode 1 or by mode 2 independent of voting preferences: 
  - $p_{m2,a,i}=p_{m2,a,i}$

Note: As we will investigate later, this condition might be needed in order to restore a rigged election to its original form.

## {{< var sec1.un6 >}}

1. Differences in the cost of voting (e.g., topology, voting rules, dropboxes etc.) across precincts $\Rightarrow$
1. Changes to the likelihood of voting $\Rightarrow$ 
1. Simplify so that changes in the likelihood of voting only impact Mode 2 $\Rightarrow$ 
1. Higher $\zeta_i$ $\Rightarrow$
1. Which changes the total election outcome through $\alpha_i=\frac{x_i+\zeta_i y_i}{\zeta_i+x_i}$

---

The impact on the election outcome, however, would be small or close to non-existens if the conditions
of the existance of a *trivial* Manifold occuranse related to normal form are not satisfied:

- Proportion of mail-in-votes to election day vote is already high (Aggregate convergence lemma)
  - Note: $\lim_{\zeta_i \rightarrow \infty} \alpha_i=\frac{x_i+\zeta_i y_i}{\zeta_i+x_i}=y_i$
- Correlation between $x$ and $y$ is very high (Twixt lemma)
  - Note: If correlation is 1, we can set $x_i=\rho_s y_i$ so that $\alpha_i(y_i,\zeta_i)=\frac{\rho_s y_i+\zeta_i y_i}{\zeta_i+x_i}$
- Low variance to $\zeta$ 
  - Note: Invariance across precincts would imply that $\zeta_i=\zeta$, so that election outcome only determined by variations in two parameters: $\alpha_i(y_i,x_i)=\frac{x_i+\zeta y_i}{\zeta+x_i}$

If none of this three cases are satisfied, variation in $\zeta_i$ should explain part of the election result. But how much should we expect within a county (2%, 5%, 10%)?

[Application 1: Election process simulator]({{< var shn1 >}})

## {{< var sec1.un7 >}}

1. Differenc in the cost of voting (e.g., topology, voting rules, dropboxes, socio economics etc.) across precincts $\Rightarrow$ 
1. Make changes to the likelihood of voting $\Rightarrow$ 
1. Simplify so that changes in the likelihood of voting only impact mode 2 $\Rightarrow$ 
1. Higher $\zeta_i$ $\Rightarrow$
1. If new and earlier sample not representative of each other, higher $\Gamma_i$ $\Rightarrow$
1. Which changes the total election outcome through $\alpha_i=\frac{g_i+\zeta_i h_i}{\zeta_i+1}$

---

The impact on the election outcome, however, would be small or close to non-existens if the conditions
of the existance of a *trivial* Manifold related to opposition form are not satisfied:

- Proportion of the hybrid ballot boxes is already high (Aggregate convergence lemma)
  - Note: $\lim_{\Gamma_i \rightarrow \infty} \alpha_i=\frac{g_i+\zeta_i h_i}{\zeta_i+g_i}=y_i$
- Correlation between $g$ and $h$ is very high (Twixt lemma)
  - Note: If correlation is 1, we can set $g_i=\rho_h h_i$ so that $\alpha_i(h_i,\Gamma_i)=\frac{\rho_h h_i+\Gamma_i h_i}{\Gamma_i+x_i}$
- Low variance to $\Gamma_i$ 
  - Note: Invariance across precincts would imply that $\Gamma_i=\Gamma$, so that election outcome only determined by variations in two parameters: $\alpha_i=\frac{g_i+\Gamma h_i}{h_i}$

If none of this three cases are satisfied, variation in $\Gamma_i$ should explain part of the election result. But how much should we expect within a county (2%, 5%, 10%)?

[Application 1: Election process simulator]({{< var shn1 >}})

# {{< var sec2.nmn >}}
## {{< var sec2.un1 >}}
- We have identified three ways of counting
  - *Standard form* ( $x$, $y$, $\zeta$, $\alpha$, $\lambda$ )
  - *Hybrid form* ( $g$, $h$, $\alpha$, $\Gamma$, $\Omega$ )
  - *Opposition form* ( $n$, $m$, $\xi$, $\Omega$, $\lambda$ )
-  Each form can (throught its parametric tautologies) determine the 5 parameters that have been listed.
  -  We are free to *predetermine* **3** of them
  -  The remaining **2** must be *backsolved*, this based on information from the first three parameters.
  -  Selecting **3** parameters from a set of **5** implies **10** different combinations

## {{< var sec2.un2 >}}
::: {.callout-caution collapse="true"}
**20 laws and 40 isometris**

```{r, echo=F}
# Load necessary library
library(tibble)

# Create the data.frame
laws_df <- tribble(
  ~Law_Number, ~North_vs_South, ~West_vs_East, ~Diagonal_vs_Diagonal,
  "First Law", "$x_{1}=\\alpha_{1}+\\zeta(\\alpha_{1}-y_{1})$", "$g_{1}=\\alpha_{1}+\\gamma(\\alpha_{1}-h_{1})$", "$m_{1}=\\Omega_{1}+\\xi(\\Omega_{1}-n_{1})$",
  "Second Law", "$x_{1}=\\lambda_{1}+\\zeta(\\lambda_{1}-y_{2})$", "$g_{1}=\\Omega_{1}+\\gamma(\\Omega_{1}-h_{2})$", "$m_{1}=\\lambda_{1}+\\xi(\\lambda_{1}-n_{2})$",
  "Third Law", "$x_{1}=\\frac{\\alpha_{1}y_{2}-\\lambda_{1}y_{1}}{(\\alpha_{1}-\\lambda_{1})-(y_{1}-y_{2})}$", "$g_{1}=\\frac{\\alpha_{1}h_{2}-\\Omega_{1}h_{1}}{(\\alpha_{1}-\\Omega_{1})-(h_{1}-h_{2})}$", "$m_{1}=\\frac{\\Omega_{1}n_{2}-\\lambda_{1}n_{1}}{(\\Omega_{1}-\\lambda_{1})-(n_{1}-n_{2})}$",
  "Fourth Law", "$x_{1}=\\frac{\\lambda_{1}+\\alpha_{1}-\\Omega_{2}}{2\\Omega_{1}}$", "$g_{1}=\\frac{\\Omega_{1}+\\alpha_{1}-\\lambda_{2}}{2\\lambda_{1}}$", "$m_{1}=\\frac{\\lambda_{1}+\\Omega_{1}-\\alpha_{2}}{2\\alpha_{1}}$",
  "Fifth Law", "$y_{1}=\\alpha_{1}-\\zeta^{-1}(\\alpha_{1}-x_{1})$", "$h_{1}=\\alpha_{1}-\\gamma^{-1}(\\alpha_{1}-g_{1})$", "$n_{1}=\\Omega_{1}-\\xi^{-1}(\\Omega_{1}-m_{1})$",
  "Sixth Law", "$y_{1}=\\lambda_{2}-\\zeta^{-1}(\\lambda_{1}-x_{1})$", "$h_{1}=\\Omega_{2}-\\gamma^{-1}(\\Omega_{1}-g_{1})$", "$n_{1}=\\lambda_{2}-\\xi^{-1}(\\lambda_{1}-m_{1})$",
  "Seventh Law", "$y_{1}=\\frac{x_{1}\\lambda_{2}-x_{2}\\alpha_{1}}{(\\lambda_{2}-\\alpha_{1})-(x_{2}-x_{1})}$", "$h_{1}=\\frac{g_{1}\\Omega_{2}-g_{2}\\alpha_{1}}{(\\Omega_{2}-\\alpha_{1})-(g_{2}-g_{1})}$", "$n_{1}=\\frac{m_{1}\\lambda_{2}-m_{2}\\Omega_{1}}{(\\lambda_{2}-\\Omega_{1})-(m_{2}-m_{1})}$",
  "Eighth Law", "$y_{1}=\\frac{\\lambda_{2}+\\alpha_{1}-\\Omega_{1}}{2\\Omega_{2}}$", "$h_{1}=\\frac{\\Omega_{2}+\\alpha_{1}-\\lambda_{1}}{2\\lambda_{2}}$", "$n_{1}=\\frac{\\lambda_{2}+\\Omega_{1}-\\alpha_{1}}{2\\alpha_{2}}$",
  "Ninth Law", "$\\alpha_{1}=x_{1}\\Omega_{1}+\\Omega_{2}y_{1}$", "$\\alpha_{1}=g_{1}\\lambda_{1}+\\lambda_{2}h_{1}$", "$\\Omega_{1}=m_{1}\\alpha_{1}+\\alpha_{2}n_{1}$",
  "Tenth Law", "$\\alpha_{1}=\\Omega_{1}(x_{1}-x_{2})+\\lambda_{2}$", "$\\alpha_{1}=\\lambda_{1}(g_{1}-g_{2})+\\Omega_{2}$", "$\\Omega_{1}=\\alpha_{1}(m_{1}-m_{2})+\\lambda_{2}$",
  "Eleventh Law", "$\\alpha_{1}=\\Omega_{2}(y_{1}-y_{2})+\\lambda_{1}$", "$\\alpha_{1}=\\lambda_{2}(h_{1}-h_{2})+\\Omega_{1}$", "$\\Omega_{1}=\\alpha_{2}(n_{1}-n_{2})+\\lambda_{1}$",
  "Twelfth Law", "$\\alpha_{1}=\\frac{x_{1}(y_{2}-y_{1})-\\lambda_{1}(x_{1}-y_{1})}{y_{2}-x_{1}}$", "$\\alpha_{1}=\\frac{g_{1}(h_{2}-h_{1})-\\Omega_{1}(g_{1}-h_{1})}{h_{2}-g_{1}}$", "$\\Omega_{1}=\\frac{m_{1}(n_{2}-n_{1})-\\lambda_{1}(m_{1}-n_{1})}{n_{2}-m_{1}}$",
  "Thirteenth Law", "$\\lambda_{1}=x_{1}\\Omega_{1}+\\Omega_{2}y_{2}$", "$\\Omega_{1}=g_{1}\\lambda_{1}+\\lambda_{2}h_{2}$", "$\\lambda_{1}=m_{1}\\alpha_{1}+\\alpha_{2}n_{2}$",
  "Fourteenth Law", "$\\lambda_{1}=\\Omega_{1}(x_{1}-x_{2})+\\alpha_{2}$", "$\\Omega_{1}=\\lambda_{1}(g_{1}-g_{2})+\\alpha_{2}$", "$\\lambda_{1}=\\alpha_{1}(m_{1}-m_{2})+\\Omega_{2}$",
  "Fifteenth Law", "$\\lambda_{1}=\\frac{\\alpha_{1}(x_{1}-y_{2})-x_{1}(y_{1}-y_{2})}{x_{1}-y_{1}}$", "$\\Omega_{1}=\\frac{\\alpha_{1}(g_{1}-h_{2})-g_{1}(h_{1}-h_{2})}{g_{1}-h_{1}}$", "$\\lambda_{1}=\\frac{\\Omega_{1}(m_{1}-n_{2})-m_{1}(n_{1}-n_{2})}{m_{1}-n_{1}}$",
  "Sixteenth Law", "$\\lambda_{1}=\\Omega_{2}(y_{2}-y_{1})+\\alpha_{1}$", "$\\Omega_{1}=\\lambda_{2}(h_{2}-h_{1})+\\alpha_{1}$", "$\\lambda_{1}=\\alpha_{2}(n_{2}-n_{1})+\\Omega_{1}$",
  "Seventeenth Law", "$\\zeta=\\frac{x_{1}-\\alpha_{1}}{\\alpha_{1}-y_{1}};\\Omega_{1}=\\frac{y_{1}-\\alpha_{1}}{y_{1}-x_{1}}$", "$\\gamma=\\frac{g_{1}-\\alpha_{1}}{\\alpha_{1}-h_{1}};\\lambda_{1}=\\frac{h_{1}-\\alpha_{1}}{h_{1}-g_{1}}$", "$\\xi=\\frac{m_{1}-\\Omega_{1}}{\\Omega_{1}-n_{1}};\\alpha_{1}=\\frac{n_{1}-\\Omega_{1}}{n_{1}-m_{1}}$",
  "Eighteenth Law", "$\\Omega_{1}=\\frac{\\lambda_{2}-\\alpha_{1}}{x_{2}-x_{1}}=\\frac{\\alpha_{2}-\\lambda_{1}}{x_{2}-x_{1}}$", "$\\gamma_{1}=\\frac{\\Omega_{2}-\\alpha_{1}}{g_{2}-g_{1}}=\\frac{\\alpha_{2}-\\Omega_{1}}{g_{2}-g_{1}}$", "$\\alpha_{1}=\\frac{\\lambda_{2}-\\Omega_{1}}{m_{2}-m_{1}}=\\frac{\\Omega_{2}-\\lambda_{1}}{m_{2}-m_{1}}$",
  "Nineteenth Law", "$\\zeta=\\frac{x_{1}-\\lambda_{1}}{\\lambda_{1}-y_{2}}; \\Omega_{1}=\\frac{y_{2}-\\lambda_{1}}{y_{2}-x_{1}}$", "$\\gamma=\\frac{g_{1}-\\Omega_{1}}{\\Omega_{1}-h_{2}}; \\lambda_{1}=\\frac{h_{2}-\\Omega_{1}}{h_{2}-g_{1}}$", "$\\xi=\\frac{m_{1}-\\lambda_{1}}{\\lambda_{1}-n_{2}}; \\alpha_{1}=\\frac{n_{2}-\\lambda_{1}}{n_{2}-m_{1}}$",
  "Twentieth Law", "$\\zeta=\\frac{\\lambda_{1}-\\alpha_{1}}{(y_{2}-y_{1})+(\\alpha_{1}-\\lambda_{1})}$", "$\\gamma=\\frac{\\Omega_{1}-\\alpha_{1}}{(h_{2}-h_{1})+(\\alpha_{1}-\\Omega_{1})}$", "$\\xi=\\frac{\\lambda_{1}-\\Omega_{1}}{(n_{2}-n_{1})+(\\Omega_{1}-\\lambda_{1})}$"
)
# Display the data.frame
kable(laws_df, format = "markdown", escape = FALSE) %>% kableExtra::kable_paper(full_width = F) %>% kableExtra::kable_styling(font_size=8)
```

:::

---

::: {.callout-caution collapse="true"}
## Proportions

```{r, echo=F}
library(kableExtra)
library(tibble)
new_tbl <- data.frame(
  V1 = c("North Ratio", "West Ratio", "Northwest Ratio", "Diagonal Aggregate", "Diagonal Complement", "West Aggregate", "East Aggregate", "North Aggregate", "South Aggregate"),
  V2 = c("South Ratio", "East Ratio", "Northeast Ratio", "Diagonal Proportion", "Diagonal Inverse", "East to West Proportion", "West to East Proportion", "South to North Proportion", "North to South Proportion"),
  V3 = c("North Complement", "West Complement", "Southeast Ratio", "1st Alpha Identity", "2nd Alpha Identity", "1st Lambda Identity", "2nd Lambda Identity", "1st Omega Identity", "2nd Omega Identity"),
  V4 = c("South Complement", "East Complement", "Southwest Ratio", "Xi Identity", "Inverse Xi Identity", "Gamma Identity", "Inverse Gamma Identity", "Zeta Identity", "Inverse Zeta Identity"),
  Equation1 = c("$x_{1} = \\frac{s}{s+t}$", "$g_{1} = \\frac{s}{s+v}$", "$m_{1} = \\frac{s}{s+u}$", "$\\alpha_{1} = \\frac{s+u}{(s+u)+(t+v)}$", "$\\alpha_{2} = \\frac{t+v}{(s+u)+(t+v)}$", "$\\lambda_{1} = \\frac{s+v}{(s+v)+(u+t)}$", "$\\lambda_{2} = \\frac{u+t}{(s+v)+(u+t)}$", "$\\Omega_{1} = \\frac{s+t}{(s+t)+(u+v)}$", "$\\Omega_{2} = \\frac{u+v}{(s+t)+(u+v)}$"),
  Equation2 = c("$y_{1} = \\frac{u}{u+v}$", "$h_{1} = \\frac{u}{u+t}$", "$n_{1} = \\frac{t}{t+v}$", "$\\xi = \\frac{t+v}{s+u}$", "$\\xi^{-1} = \\frac{s+u}{t+v}$", "$\\gamma = \\frac{u+t}{s+v}$", "$\\gamma^{-1} = \\frac{s+v}{u+t}$", "$\\zeta = \\frac{u+v}{s+t}$", "$\\zeta^{-1} = \\frac{s+t}{u+v}$"),
  Equation3 = c("$x_{2} = (1-x_{1}) = \\frac{t}{s+t}$", "$g_{2} = (1-g_{1}) = \\frac{v}{s+v}$", "$m_{2} = (1-m_{1}) = \\frac{u}{s+u}$", "$\\alpha_{1} = (\\xi+1)^{-1}$", "$\\alpha_{2} = (\\xi^{-1}+1)^{-1}$", "$\\lambda_{1} = (\\gamma+1)^{-1}$", "$\\lambda_{2} = (\\gamma^{-1}+1)^{-1}$", "$\\Omega_{1} = (\\zeta+1)^{-1}$", "$\\Omega_{2} = (\\zeta^{-1}+1)^{-1}$"),
  Equation4 = c("$y_{2} = (1-y_{1}) = \\frac{v}{u+v}$", "$h_{2} = (1-h_{1}) = \\frac{t}{u+t}$", "$n_{2} = (1-n_{1}) = \\frac{v}{t+v}$", "$\\xi = \\frac{1-\\alpha_{1}}{\\alpha_{1}} = \\frac{\\alpha_{2}}{\\alpha_{1}}$", "$\\xi^{-1} = \\frac{1-\\alpha_{2}}{\\alpha_{2}} = \\frac{\\alpha_{1}}{\\alpha_{2}}$", "$\\gamma = \\frac{1-\\lambda_{1}}{\\lambda_{1}} = \\frac{\\lambda_{2}}{\\lambda_{1}}$", "$\\gamma^{-1} = \\frac{1-\\lambda_{2}}{\\lambda_{2}} = \\frac{\\lambda_{1}}{\\lambda_{2}}$", "$\\zeta = \\frac{1-\\Omega_{1}}{\\Omega_{1}} = \\frac{\\Omega_{2}}{\\Omega_{1}}$", "$\\zeta^{-1} = \\frac{1-\\Omega_{2}}{\\Omega_{2}} = \\frac{\\Omega_{1}}{\\Omega_{2}}$")
)

# Extract the columns
V1 <- new_tbl$V1
V2 <- new_tbl$V2
V3 <- new_tbl$V3
V4 <- new_tbl$V4
Equation1 <- new_tbl$Equation1
Equation2 <- new_tbl$Equation2
Equation3 <- new_tbl$Equation3
Equation4 <- new_tbl$Equation4

# Combine the columns
combined <- data.frame(
  V1 = matrix(c(rbind(V1, Equation1)), ncol = 1),
  V2 = matrix(c(rbind(V2, Equation2)), ncol = 1),
  V3 = matrix(c(rbind(V3, Equation3)), ncol = 1),
  V4 = matrix(c(rbind(V4, Equation4)), ncol = 1)
)
kable(combined, format = "markdown", escape = FALSE) %>% kableExtra::kable_paper(full_width = F) %>% kableExtra::kable_styling(font_size=8)
```

:::
## {{< var sec2.un3 >}}