### 2024-07-13 analyzer.qmd
sommore

---
title: ""
output: github_document
always_allow_html: true
execute: 
  cache: false
webr:
  show-startup-message: false      # Display status of webR initialization
  show-header-message: false      # Check to see if COOP&COEP headers are set for speed.
  packages: ['dplyr','data.table','ggplot2','combinat','plotly','htmltools','broom'] # Pre-install dependencies
  cell-options:
    autorun: true
filters:
  - panelize
  - webr
  - pyodide
---

```{webr-r}
#| echo: fenced
#| eval: true
webr::install("ManifoldDestinyWASMP", repos = "https://lotariohw26.github.io/MD_WASMC")
webr::install("ManifoldDestinyWASMD", repos = "https://lotariohw26.github.io/MD_WASMC")
library(ManifoldDestinyWASMP)
library(ManifoldDestinyWASMD)
ls(package:ManifoldDestinyWASMP)
data(package='ManifoldDestinyWASMD')
dfl <- ManifoldDestinyWASMD::yout
```

```{webr-r}
apps <- basc
adat <- apps[[1]]
amet <- apps[[2]]
cogr <- Countinggraphs(apps[[1]])
#selo <- seloutput(selreport(apps))
#copl <- cogr$polyc[[1]][[1]]
cogr$sdfc
```