### 2024-07-13 rigging.qmd

---
title: ""
output: github_document
always_allow_html: true
execute: 
  cache: false
webr:
  show-startup-message: false      # Display status of webR initialization
  show-header-message: false      # Check to see if COOP&COEP headers are set for speed.
  packages: ['dplyr','data.table','ggplot2','combinat','plotly','htmltools','broom'] # Pre-install dependencies
  cell-options:
    autorun: true
filters:
  - panelize
  - webr
  - pyodide
---

```{webr-r}
#| echo: fenced
#| eval: true
webr::install("ManifoldDestinyWASMP", repos = "https://lotariohw26.github.io/MD_WASMC")
webr::install("ManifoldDestinyWASMD", repos = "https://lotariohw26.github.io/MD_WASMC")
library(ManifoldDestinyWASMP)
library(ManifoldDestinyWASMD)
ls(package:ManifoldDestinyWASMP)
data(package='ManifoldDestinyWASMD')
dfl <- ManifoldDestinyWASMD::yout
```

```{webr-r}
###### Normal 
# Simulation
perv <- (function(x){data.frame(P=seq(1,x),RV=as.integer(rnorm(x,100,30)))})(100)
prow <- c(m=0.51,s=0.10)
proa <- c(vdm=0.7,mdm=0.4,vds=0.10,mds=0.10)
prob <- c(vdm=0.5,mdm=0.6,vds=0.10,mds=0.10)
ztec <- c(0,1)	
basc <- list(df=ballcastsim(perv,prow,proa,prob,ztec),list(fr=1,eq="alpha=k0+k1*x+k2*y",va='y',prg=list(cnd=1)))
# Rigged election
cogr <- Countinggraphs(basc[[1]])
copl <- cogr$polyc[[1]][[1]]
plfc <- c(0.20,0.20,0.20)
ex1s <- list(frm=1, pre=c("alpha","x","y"), end=c("zeta","Omega","lamda"),
	     stuv=c("S","T","U","V"),
	     plnr=1,rot=list(fr=c(1,10),sr=c(4,0),tr=c(2,0)),
	     ls="(alpha-alpha_s)^2"
)
ex2s <- list(frm=1, pre=c("alpha","x","y"), end=c("zeta","Omega","lamda"),
	     stuv=c("S","T","U","V"),
	     plnr=1,rot=list(fr=c(1,10),sr=c(4,0),tr=c(2,0)),
	     ls="(alpha-alpha_s)^2"
)
ex3s <- list(frm=1, pre=c("alpha","x","y"), end=c("zeta","Omega","lamda"),
	     stuv=c("S","T","U","V"),
	     plnr=1,rot=list(fr=c(1,10),sr=c(4,0),tr=c(2,0)),
	     ls="(alpha-alpha_s)^2"
)
exs <- list(ex1s,ex2s,ex3s)
rigv <- lapply(1:3, function(x) { 
  cogr$mansys(sygen=exs[[x]])
  cogr$polyc[[x]]
  cogr$setres(NULL,1)
  cogr$setres(plfc[x],1)
  cogr$manimp(init_par=c(k0=0,k1=0.60,k2=0.40),wn=c(0,0),man=T)
  exs[[1]]$frm
  rasc <- list(cogr$rdfc,list(fr=1,eq="alpha=k0+k1*x+k2*y",va='y',prg=list(cnd=1)))
})
###################################################################################################################################################################
```

```{webr-r}
appnf <- seloutput(rppnf <- selreport(basc))
appr1 <- seloutput(rppr1 <- selreport(rigv[[1]]))
appr2 <- seloutput(rppr2 <- selreport(rigv[[2]]))
appr3 <- seloutput(rppr3 <- selreport(rigv[[3]]))
penf <- polynom::polynomial(unname(rppnf[[1]]$polyc[[1]][[1]],3))
per1 <- polynom::polynomial(unname(rppr1[[1]]$polyc[[1]][[1]],3))
per2 <- polynom::polynomial(unname(rppr2[[1]]$polyc[[1]][[1]],3))
per3 <- polynom::polynomial(unname(rppr3[[1]]$polyc[[1]][[1]],3))
```

```{webr-r}
appnf[[1]]
```