### 2024-08-17 index.qmd
late

---
title: "{{< var titll >}}"
output: github_document
execute: 
  cache: false
#webr:
#  show-startup-message: true      # Display status of webR initialization
#  show-header-message: false      # Check to see if COOP&COEP headers are set for speed.
#  packages: ['ggplot2','dplyr','data.table','readr','plotly','combinat','htmltools','broom','cowplot'] 
#  cell-options:
#    autorun: true
#filters:
#  - panelize
#  - webr
---

```{r, echo=F}
# script/R/simulations.R
```

**Based upon the ideas developed by Edward K. Solomon**
![](pngs/md.png)

```{r, echo=F}
suppressMessages(ManifoldDestiny::wasmconload()) 
apps <- appn0
adat <- apps[[1]]
amet <- apps[[2]]
cob <- Countinggraphs(adat,selvar=names(adat))
sum(dplyr::select(cob$rdfci,S,T,U,V))
cob$purging(z=amet$prg$z,pri=1) 
cob$plext(2)
sum(dplyr::select(cob$rdfce,S,T,U,V))
def <- Estimation(cob$rdfce)
def$regression("alpha=k0 + k1*g + k2*h")
summary(def$regsum[[1]])
```

```{r}
def$regression("alpha=k0 + k1*g + k2*h")
summary(def$regsum[[1]])
```

```{r}
def$regression("alpha=k0 + k1*g + k2*h + k3*g**2 + k4*h**2 + k5*g*h + k6*h**3 + k7*g*h + k8*g**2*h + k9*g*h**2")
summary(def$regsum[[1]])
```

```{r, echo=F}
knitr::knit_exit()
```

```{webr-r}
#| echo: fenced
#| eval: true
webr::install("ManifoldDestinyWASMP", repos = "https://lotariohw26.github.io/MD_WASMC")
webr::install("ManifoldDestinyWASMD", repos = "https://lotariohw26.github.io/MD_WASMC")
library(ManifoldDestinyWASMP)
library(ManifoldDestinyWASMD)
ls(package:ManifoldDestinyWASMP)
data(package='ManifoldDestinyWASMD')
dfl <- ManifoldDestinyWASMD::yout
```

```{r, echo=F}
knitr::knit_exit()
```

```{webr-r}
coi <- Countingprocess(dfl[[1]])
ges <- Estimation(coi$rdfc,2)
ges$regression("alpha=k0+k1*g+k2*h")
ges$diagnostics()
broom::tidy(ges$regsum[[1]])
broom::glance(ges$regsum[[1]])
```

[Nevada](https://docs.google.com/spreadsheets/d/1gCfmuK44sUeb0Hb0plAc8VllPSjtUiEHWlJjeHq3xXY/edit?gid=0#gid=0)

**Last time updated: `r format(Sys.time(), "%Y-%m-%d %H:%M:%S")`**