### 2024-08-17 index.qmd

---
title: "{{< var titll >}}"
output: github_document
execute: 
  cache: false
webr:
  show-startup-message: true      # Display status of webR initialization
  show-header-message: false      # Check to see if COOP&COEP headers are set for speed.
  packages: ['ggplot2','dplyr','data.table','readr','plotly','combinat','htmltools','broom','cowplot'] 
  cell-options:
    autorun: true
filters:
  - panelize
  - webr
---

```{r, echo=F}
# script/R/simulations.R
```

**Based upon the ideas developed by Edward K. Solomon**
![](pngs/md.png)

**Last time updated: `r format(Sys.time(), "%Y-%m-%d %H:%M:%S")`**