### 2024-10-20 abc.qmd
cont

# Abc

[abc](https://www.arapahoevotes.gov/sites/default/files/2023-12/REDACTED_CVR_Export_20231128_0.xls)