### 2024-10-29 tabset.qmd
abc

---
title: ""
---

```{r}
#| echo: false
#| output: false
#| message: false
ManifoldDestiny::wasmconload()
source(paste0(rprojroot::find_rstudio_root_file(),"/R/wasmconverting.R"))
source(paste0(rprojroot::find_rstudio_root_file(),"/R/wasmnonverting.R"))
```

::: {.callout-caution collapse="true"}
## {{< var tabh.h0 >}}
```{r}
#| echo: false
#| output: false
apps <- apsnn
```
{{< include tab.qmd >}}
:::
::: {.callout-caution collapse="true"}
## {{< var tabh.h1 >}}
```{r}
#| echo: false
#| output: false
apps <- aprnn
```
{{< include tab.qmd >}}
:::

::: {.callout-caution collapse="true"}
## {{< var tabh.h2 >}}
```{r}
#| echo: false
#| output: false
apps <- aprnn
```
{{< include tab.qmd >}}
:::

::: {.callout-caution collapse="true"}
## {{< var tabh.h3 >}}
```{r}
#| echo: false
#| output: false
apps <- aprnn
```
{{< include tab.qmd >}}
:::
::: {.callout-caution collapse="true"}
## {{< var sec6.un1 >}}
```{r}
#| echo: false
#| output: false
apps <- apn1n
```
{{< include tab.qmd >}}
:::
::: {.callout-caution collapse="true"}
## {{< var sec6.un2 >}}
```{r}
#| echo: false
#| output: false
apps <- apn1n
```
{{< include tab.qmd >}}
:::
```{r}
knitr::knit_exit()
```
::: {.callout-caution collapse="true"}
## {{< var sec6.un3 >}}
```{r}
#| echo: false
#| output: false
apps <- appn3
```
{{< include tab.qmd >}}
:::

::: {.callout-caution collapse="true"}
## {{< var sec6.un4 >}}
```{r}
#| echo: false
#| output: false
apps <- appn4
```
{{< include tab.qmd >}}
:::