### 2024-11-15 docs/analyzer.qmd
rec

---
title: ""
output: github_document
always_allow_html: true
execute: 
  cache: false
webr:
  show-startup-message: false      # Display status of webR initialization
  show-header-message: false      # Check to see if COOP&COEP headers are set for speed.
  packages: ['dplyr','data.table','ggplot2','combinat','plotly','htmltools','broom'] # Pre-install dependencies
  cell-options:
    autorun: true
filters:
  - panelize
  - webr
  - pyodide
---

```{webr-r}
#| echo: fenced
#| eval: true
webr::install("ManifoldDestinyWASMP", repos = "https://lotariohw26.github.io/MD_WASMC")
webr::install("ManifoldDestinyWASMD", repos = "https://lotariohw26.github.io/MD_WASMC")
library(ManifoldDestinyWASMP)
library(ManifoldDestinyWASMD)
ls(package:ManifoldDestinyWASMP)
data(package='ManifoldDestinyWASMD')
```

```{webr-r}
#apps <- appn0
#adat <- apps[[1]]
#amet <- apps[[2]]
#cob <- Countinggraphs(adat)
#cob$purging(z=0) 
#def <- Estimation(cob$rdfce)
```

```{webr-r}
def$regression("alpha=k0 + k1*g + k2*h")
summary(def$regsum[[1]])
```

```{webr-r}
def$regression("alpha=k0 + k1*g + k2*h + k3*g**2 + k4*h**2 + k5*g*h + k6*h**3 + k7*g*h + k8*g**2*h + k9*g*h**2")
summary(def$regsum[[1]])
```